---
title: "Fit GLMMs"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit GLMMs}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
knitr::opts_chunk$set(progress = TRUE, verbose = TRUE)
knitr::opts_chunk$set(message = FALSE) 
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 7,
  fig.height = 7,
  fig.align = "center"
)
```

Load libraries

```{r libraries, echo=TRUE, warning=FALSE}
library("readxl")
library("tidyverse")
library("dplyr")
library("glmmTMB")
library("DHARMa")
library("ggeffects")
library("here")
library("gridExtra")
library("rainy")
library("lubridate")
library("ggpubr")
```

### Import data

```{r import-data}
# event data includes dates and time out and removed, exposure hours and 
# assessment date

event_data <- read_csv(system.file(
  "extdata",
  "exposure_events.csv",
  package = "rainy",
  mustWork = TRUE
)) %>%
  mutate(
    date_time_out = dmy_hm(date_time_out),
    date_time_removed = dmy_hm(date_time_removed),
    exposed = interval(date_time_out, date_time_removed)
  )

lesion_data <- read_csv(system.file(
  "extdata",
  "lesion_data.csv",
  package = "rainy",
  mustWork = TRUE
)) %>%
  pivot_longer(cols = plant_1:plant_3,
               names_to = "plant",
               values_to = "lesions") %>% 
  group_by(rainfall_event, pot) %>% 
  summarise(total_lesions = sum(lesions, na.rm = TRUE),
         mean_lesions = round(mean(lesions), 1))
```

## Create object for analysis

Lastly we import the weather data and then:

 * filter for only the dates/times with plants deployed,
 * summarise by rainfall event, and then
 * join with the lesion data.

```{r create_dat}
dat <- read_csv(system.file(
  "extdata",
  "weather_data.csv",
  package = "rainy",
  mustWork = TRUE
)) %>%
  mutate(
    Date_time = round_date(dmy_hms(paste(Date, Time)), "minute"),
    rain_duration = as.integer(`TOTAL Rain Gauge mm` > 0),
    rainfall_event = case_when(
      Date_time %within% event_data[1, "exposed"] ~ 1,
      Date_time %within% event_data[2, "exposed"] ~ 2,
      Date_time %within% event_data[3, "exposed"] ~ 3,
      Date_time %within% event_data[4, "exposed"] ~ 4,
      Date_time %within% event_data[5, "exposed"] ~ 5,
      Date_time %within% event_data[6, "exposed"] ~ 6,
      Date_time %within% event_data[7, "exposed"] ~ 7
    )
  ) %>%
  filter(
    Date_time %within% event_data[1, "exposed"] |
      Date_time %within% event_data[2, "exposed"] |
      Date_time %within% event_data[3, "exposed"] |
      Date_time %within% event_data[4, "exposed"] |
      Date_time %within% event_data[5, "exposed"] |
      Date_time %within% event_data[6, "exposed"] |
      Date_time %within% event_data[7, "exposed"]
  ) %>%
  group_by(rainfall_event) %>%
  summarise(
    wind_direction = round(circular.averaging(`AVERAGE Wind Direction Degs`), 1),
    total_rain = sum(`TOTAL Rain Gauge mm`),
    rain_duration = round(sum(rain_duration / 6), 1),
    temp = round(mean(`AVERAGE Air Temperature DegC`), 1),
    relative_humidity = round(mean(`AVERAGE Relative Humidity %`), 1),
    wind_speed = round(mean(`AVERAGE Wind Speed km/h`) * 0.277778, 1) # convert to m/s
  ) %>%
  left_join(lesion_data, by = "rainfall_event") %>%
  mutate(rainfall_event = as.factor(rainfall_event))
```

Use `set.seed()` for reproducibility purposes

```{r set-seed, echo=TRUE}
set.seed(42)
```

### Model_1 (All variables)

Include all variables with rainfall event as a random effect.

```{r mod1}
mod1 <-
  glmmTMB(
    total_lesions ~ total_rain +
      rain_duration +
      relative_humidity +
      temp +
      wind_speed +
      wind_direction +
      (1 | rainfall_event),
    family = nbinom1,
    data = dat
  )

summary(mod1)
```

The model summary suggest that there are not enough degrees of freedom to fit a complicated model.
We can drop relative humidity and temperature because we directly transferred plants to a controlled environment with 20 degree Celsius temperature and 100 % RH

### Model_2

Drop relative humidity and temperature 
```{r mod2}
mod2 <-
  glmmTMB(
    total_lesions ~  total_rain +
      rain_duration +
      wind_speed +
      wind_direction + (1 | rainfall_event),
    family = nbinom1,
    data = dat
  )

summary(mod2)
```

We can see that model has been successful converged. But getting a warning message

### Model_3

Remove wind speed as it is not significant 

```{r fit-mod-3}
mod3 <-
  glmmTMB(
    total_lesions ~  total_rain +
      rain_duration +
      wind_direction + (1 | rainfall_event),
    family = nbinom1,
    data = dat
  )

summary(mod3)
```
Again getting a warning message.
And the negative impact of increasing rain duration and wind direction does not make any biological sense 

### Model_4

Replace wind direction with wind speed if the warning message could be removed

```{r fit-mod4}
mod4 <-
  glmmTMB(
    total_lesions ~
      total_rain +
      rain_duration +
      wind_speed +
      (1 | rainfall_event),
    family = nbinom1,
    data = dat
  )

summary(mod4)
```

The warning message has gone now. The reason is that wind direction did not vary much during different rainfall events, whereas wind speed varied. We can go with `mod4`.

## Best fit model `mod4` diagnostics

### Check over-dispersion

Check for over-dispersion using custom function created by Ben Bolker.
Here is the [link](http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#overdispersion).

```{r check-over-dispersion, warning=FALSE}
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model, type = "pearson")
  Pearson.chisq <- sum(rp ^ 2)
  prat <- Pearson.chisq / rdf
  pval <- pchisq(Pearson.chisq, df = rdf, lower.tail = FALSE)
  c(
    chisq = Pearson.chisq,
    ratio = prat,
    rdf = rdf,
    p = pval
  )
}
overdisp_fun(mod4)
```

The larger `p value` shows that there is no significant over-dispersion.

### Simulate residuals


```{r simulate-residuals, warning=FALSE}
simulateResiduals(mod4, plot = T, quantreg = T)
```

Very close observed and predicted values 

### Check zero-inflation

```{r check-zero-inflation, warning=FALSE}
testZeroInflation(mod4)
```

No zero inflation detected

### Check overall performance

```{r over-all-performance check, warning=FALSE}
performance::check_model(mod4, panel = FALSE)
```

The model fit non-linear model data distribution assumption, close observed vs. fitted values and there is no significant collinearity between predictors.

## Plot best fit model

```{r best-fit-plot, warning=FALSE}
f1 <-
  plot(ggpredict(mod4, "total_rain")) + 
  theme_pubclean(base_family = "Arial")
f2 <-
  plot(ggpredict(mod4 , "rain_duration")) + 
  theme_pubclean(base_family = "Arial")
f3 <-
  plot(ggpredict(mod4, "wind_speed")) + 
  theme_pubclean(base_family = "Arial")
fig_5 <- grid.arrange(f1, f2, f3)

fig_5

```

```{r save-fig-5, warning=FALSE}
ggsave(
   here("man", "figures/fig_4.png"),
   plot = fig_5,
   width = 9,
   height = 9,
   units = "in",
   dpi = 600
)

ggsave(
   here("man", "figures/fig_4.eps"),
   plot = fig_5,
   width = 9,
   height = 9,
   units = "in",
   device =  cairo_ps
)
```

## Colophon

```{r sessioninfo}
sessionInfo()
```
