---
title: "Fit GLMMs"
date: "29/04/2021"
output:
   rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit GLMMs}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
knitr::opts_chunk$set(progress = TRUE, verbose = TRUE)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 7,
  fig.height = 7,
  fig.align = "center"
)
```

Load libraries

```{r libraries, message=FALSE, echo=TRUE, warning=FALSE}
library("readxl")
library("tidyverse")
library("dplyr")
library("glmmTMB")
library("DHARMa")
library("ggeffects")
library("here")
library("gridExtra")
library("rainy")
```

### Import data

```{r import-data}
dat <-
  read_excel(
    system.file("extdata",
                "lesion_weather_data.xlsx",
                package = "rainy"),
    sheet = 1,
    na = "NA"
  ) %>%
  group_by(rainfall_event) %>%
  mutate(wind_direction = circular.averaging(wind_direction)) %>%
  mutate(rainfall_event = as.factor(rainfall_event)) %>%
  mutate(total_lesions = as.integer(total_lesions)) 
```

Use `set.seed()` for reproducibility purposes

```{r set-seed, echo=TRUE}
set.seed(42)
```

### Model_1 (All variables)

Include all variables with rainfall event as a random effect.

```{r fit-mod2}
mod1 <-
  glmmTMB(
    total_lesions ~ total_rain +
      rain_duration +
      relative_humidity +
      temp +
      wind_speed +
      wind_direction +
      (1 | rainfall_event),
    family = nbinom1,
    data = dat
  )

summary(mod1)
```

The model summary suggest that there are not enough degrees of freedom to fit a complicated model.
We can drop relative humidity and temperature because we directly transferred plants to a controlled environment with 20 degree Celsius temperature and 100 % RH

### Model_2

Drop relative humidity and temperature 
```{r}
mod2 <-
  glmmTMB(
    total_lesions ~  total_rain +
      rain_duration +
      wind_speed +
      wind_direction + (1 | rainfall_event),
    family = nbinom1,
    data = dat
  )

summary(mod2)
```

We can see that model has been successful converged. But getting a warning message

### Model_3

Remove wind speed as it is not significant 

```{r fit-mod-3}
mod3 <-
  glmmTMB(
    total_lesions ~  total_rain +
      rain_duration +
      wind_direction + (1 | rainfall_event),
    family = nbinom1,
    data = dat
  )

summary(mod3)
```
Again getting a warning message.
And the negative impact of increasing rain duration and wind direction does not make any biological sense 

### Model_4

Replace wind direction with wind speed if the warning message could be removed

```{r fit-mod4}
mod4 <-
  glmmTMB(
    total_lesions ~  total_rain +
      rain_duration +
      wind_speed + (1 | rainfall_event),
    family = nbinom1,
    data = dat
  )

summary(mod4)
```

The warning message has gone now. The reason is that wind direction did not vary much during different rainfall events, whereas wind speed varied. We can go with `mod4`

## Best fit model `mod4` diagnostics

### Check over-dispersion

Check for over-dispersion using custom function created by Ben Bolker. Here is the  [link](http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#overdispersion)

```{r check-over-dispersion}
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model, type = "pearson")
  Pearson.chisq <- sum(rp ^ 2)
  prat <- Pearson.chisq / rdf
  pval <- pchisq(Pearson.chisq, df = rdf, lower.tail = FALSE)
  c(
    chisq = Pearson.chisq,
    ratio = prat,
    rdf = rdf,
    p = pval
  )
}
overdisp_fun(mod4)
```

The larger `p value` shows that there is no significant over-dispersion 

### Simulate residuals


```{r simulate-residuals}
simulateResiduals(mod4, plot = T, quantreg = T)
```

Very close observed and predicted values 

## Plot best fit model

```{r}
f1 <-
  plot(ggpredict(mod4, "total_rain")) + 
  theme_pubclean(base_family = "Arial")
f2 <-
  plot(ggpredict(mod4 , "rain_duration")) + 
  theme_pubclean(base_family = "Arial")
f3 <-
  plot(ggpredict(mod4, "wind_speed")) + 
  theme_pubclean(base_family = "Arial")
fig_5 <- grid.arrange(f1, f2, f3)

fig_5

```


```{r}
ggsave(
   here("man", "figures/fig_5.png"),
   plot = fig_5,
   width = 9,
   height = 9,
   units = "in",
   dpi = 600
)

ggsave(
   here("man", "figures/fig_5.eps"),
   plot = fig_5,
   width = 9,
   height = 9,
   units = "in",
   device =  cairo_ps
)
```



