---
title: "Data visualisation"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data visualisation}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
knitr::opts_chunk$set(progress = TRUE, verbose = TRUE)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 9,
  fig.height = 9,
  fig.align = "center"
)
```

```{r message=FALSE, echo=TRUE, warning=FALSE}
library("grDevices")
library("lubridate")
library("tidyverse")
library("clifro")
library("viridis")
library("showtext")
library("here")
library("ggpubr")
library("extrafont")
library("rainy")
library("tidyr")
library("readxl")
library("kableExtra")

theme_set(theme_pubclean(base_family = "Arial", base_size = 15))
```

```{r fonts, include=FALSE, message=FALSE, eval=FALSE, message=FALSE}
# Note that GhostScript needs to be installed at the system level for the PS files to be generated.
# MacOS users can use `brew install ghostscript`
# 
# Windows users can follow these directions:
# 1.	Go to the GhostScript website (https://www.ghostscript.com/download/gsdnld.html)
# 2.	Download the windows installer suitable for your machine
# 3.	Run the installer file which you downloaded and follow the prompts
# 4.	After running the installer click the windows "Start" button and type "Edit environment variables for your account" and open
# 5.	In the tab 'Advanced' click the button at the bottom 'Environment Variables...'
# 6.	Under 'System variables' find the variable 'Path', select 'Path' and click the 'Edit' button
# 7. 	Select a new line and copy the Ghostscript 'bin' folder location into the field.
# 7.1	If you installed ghostscript to the default folder location; then the folder location will likely be "C:\Program Files\gs\gs9.52\bin", the version number (9.52) may differ.
# 8.	Save and exit the environmental variables window


# This chunk is then run only if knitting on new computer that the files have not been generated on
# this is necessary to embed fonts in .eps files for EJPP
library("extrafont")

if (.Platform$OS.type == "windows") {
   font_import(pattern = "arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
} else {
   font_import(pattern = "Arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
}
```


### Import data 


```{r import-data, message=FALSE}
# event data includes dates and time out and removed, exposure hours and 
# assessment date

event_data <- read_csv(system.file(
  "extdata",
  "exposure_events.csv",
  package = "rainy",
  mustWork = TRUE
)) %>%
  mutate(
    date_time_out = dmy_hm(date_time_out),
    date_time_removed = dmy_hm(date_time_removed),
    exposed = interval(date_time_out, date_time_removed)
  )

lesion_data <- read_csv(system.file(
  "extdata",
  "lesion_data.csv",
  package = "rainy",
  mustWork = TRUE
)) %>%
  pivot_longer(cols = plant_1:plant_3,
               names_to = "plant",
               values_to = "lesions") %>% 
  group_by(rainfall_event, pot) %>% 
  summarise(total_lesions = sum(lesions, na.rm = TRUE),
         mean_lesions = round(mean(lesions), 1))
```

## Create object for analysis

Lastly we import the weather data and then:

 * filter for only the dates/times with plants deployed,
 * summarise by rainfall event,
 * join with the lesion data, and then
 * create a new column with the `total_lesions` outliers filtered to use in the box plot, where outliers appear as a part of the `geom_boxplot()` but not in the `geom_jitter()`.

```{r create_dat}
dat <- read_csv(system.file(
  "extdata",
  "weather_data.csv",
  package = "rainy",
  mustWork = TRUE
)) %>%
  mutate(
    Date_time = round_date(dmy_hms(paste(Date, Time)), "minute"),
    rain_duration = as.integer(`TOTAL Rain Gauge mm` > 0),
    rainfall_event = case_when(
      Date_time %within% event_data[1, "exposed"] ~ 1,
      Date_time %within% event_data[2, "exposed"] ~ 2,
      Date_time %within% event_data[3, "exposed"] ~ 3,
      Date_time %within% event_data[4, "exposed"] ~ 4,
      Date_time %within% event_data[5, "exposed"] ~ 5,
      Date_time %within% event_data[6, "exposed"] ~ 6,
      Date_time %within% event_data[7, "exposed"] ~ 7
    )
  ) %>%
  filter(
    Date_time %within% event_data[1, "exposed"] |
      Date_time %within% event_data[2, "exposed"] |
      Date_time %within% event_data[3, "exposed"] |
      Date_time %within% event_data[4, "exposed"] |
      Date_time %within% event_data[5, "exposed"] |
      Date_time %within% event_data[6, "exposed"] |
      Date_time %within% event_data[7, "exposed"]
  ) %>%
  group_by(rainfall_event) %>%
  summarise(
    wind_direction = round(circular.averaging(`AVERAGE Wind Direction Degs`), 1),
    total_rain = sum(`TOTAL Rain Gauge mm`),
    rain_duration = round(sum(rain_duration / 6), 1),
    temp = round(mean(`AVERAGE Air Temperature DegC`), 1),
    relative_humidity = round(mean(`AVERAGE Relative Humidity %`), 1),
    wind_speed = round(mean(`AVERAGE Wind Speed km/h`) * 0.277778, 1) # convert to m/s
  ) %>%
  left_join(lesion_data, by = "rainfall_event") %>%
  mutate(
    rainfall_event = as.factor(rainfall_event))
```


Check the data.

```{r kable-dat}
kable(dat,
      format = "html",
      table.attr = "class='table table-hover'")
```

### Kernel density plot

```{r data-distribution, message=FALSE}
ggplot(dat, aes(x = mean_lesions)) +
  geom_density(fill = "steelblue", alpha = 0.7) +
  geom_rug(alpha = 0.4) +
  xlab("Total lesions per pot during each rainfall event")
```

Kernel density plots showing the shape of data distribution.
A strong peak at fewer than 25 lesions per pot was observed.

### Scatter plot

Taking the existing `dat` object, group by `rainfall_event` again and calculate the total number of lesions in 12 pots.
Plot the `total_rain` on the x-axis, `wind_speed` on the y-axis, the `total_lesions` per `rainfall_event` as a factor using colour and `rain_duration` as the point size.

```{r scatterplot, warning=FALSE}
dat %>%
  group_by(rainfall_event) %>%
  mutate(total_lesions = sum(total_lesions, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(aes(
    x = total_rain,
    y = wind_speed,
    size = rain_duration,
    colour = total_lesions
  )) +
  geom_point() +
  scale_size_continuous(range = c(3, 10),
                        breaks = c(0.7, 1.70, 3.30, 7)) +
  scale_colour_viridis_c(direction = -1, option = "E") +
  guides(
    size = guide_legend(
      title = "Rainfall duration (hour)",
      title.position = "top",
      title.hjust = 0.5
    ),
    colour = guide_legend(
      title = "Total lesions per pot",
      override.aes = list(size = 10),
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  scale_x_continuous(breaks = seq(from = 0, to = 14, by = 1),
                     limits = c(0, 14)) +
  scale_y_continuous(breaks = seq(from = 2, to = 5, by = 0.25),
                     limits = c(2.75, 4.75)) +
  labs(y = "Wind speed (m/s)",
       x = "Total rainfall (mm)") +
  theme(legend.key = element_blank(),
        plot.margin = margin(25, 25, 10, 25)) +
  coord_cartesian(clip = "off",
                  expand = FALSE)
```

```{r save-fig1, include=FALSE, eval=TRUE}
ggsave(here::here("man", "figures/Fig1.png"))
ggsave(here::here("man", "figures/Fig1.eps"), device = cairo_ps)
```

### Boxplot 

Boxplot of the total lesions for each rainfall event with total rainfall as the colour.

```{r boxplot-plot, warning=FALSE}
ggplot(
  dat,
  aes(
    x = rainfall_event,
    y = total_lesions,
    group = rainfall_event,
    colour = total_rain,
    fill = total_rain
  )
) +
  geom_boxplot(alpha = 0.5,
               outlier.shape = NA) +
  geom_jitter(width = 0.25) +
  scale_colour_viridis_c(option = "E",
                         direction = -1,
                         name = "Total rainfall (mm)") +
  scale_fill_viridis_c(option = "E",
                       direction = -1,
                       name = "Total rainfall (mm)") +
  labs(x = "Rainfall event",
       y = "Total lesions per pot") +
  guides(color = guide_colorbar(
    title.position = "top",
    title.hjust = 0.5,
    barwidth = unit(20, "lines"),
    barheight = unit(0.5, "lines")
  )) +
  theme(legend.key = element_blank(),
        plot.margin = margin(25, 25, 10, 25)) +
  coord_cartesian(clip = "off",
                  expand = FALSE)
```

```{r save-fig2, include=FALSE, eval=TRUE}
ggsave(here::here("man", "figures/Fig2.png"))
ggsave(here::here("man", "figures/Fig2.eps"), device = cairo_ps)
```

### Wind rose

Import wind speed and wind direction data.

```{r wind-rose}
wind_dat <-
  read_excel(system.file("extdata", "wind_data.xlsx", package = "rainy")) %>%
  mutate(wind_speed = wind_speed / 3.6) %>%
  mutate(wind_direction = as.numeric(wind_direction)) %>%
  mutate(wind_speed = as.numeric(wind_speed)) %>%
  mutate(rainfall_event = as.factor(rainfall_event))
```

### Plot wind rose

```{r plot-wind-rose, message=FALSE, warning=FALSE}
fig_4 <-
  with(
    wind_dat,
    windrose(
      wind_speed,
      wind_direction,
      rainfall_event,
      n_col = 3,
      legend_title = "Wind speed (m/s)"
    )
  )

fig_4 <-
  fig_4 +
  scale_fill_viridis_d(name = "Wind Speed (m/s)", direction = -1) +
  xlab("") +
  theme_pubclean(base_family = "Arial")

fig_4

```


```{r save-fig4, include=FALSE, eval=TRUE}
ggsave(here::here("man", "figures/Fig3.png"))
ggsave(here::here("man", "figures/Fig3.eps"), device = cairo_ps)
```



